\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
The thesis reveals the aggregated results of my work at the Section of Forest Growth Modelling and Computer Sciences of the Northwest German Research Institute and the Department of Forest Economics and Forest Management of the University of Göttingen. Along the way, many people made significant contribution to the presented work. Its completion would have been unlikely without the constructive help and the continuous support from my colleagues, family and friends, whom I thank herewith. In particular, I thank
\renewcommand{\labelitemi}{--}
\begin{itemize}
\item Prof. Dr. Jürgen Nagel - for the supervision of the thesis, critical revision of my workings and the freedom to explore new ideas.
\item Prof. Dr. Bernhard Möhring - for acting as co-referee, the constructive and rational support of economic problems and the opportunity to keep going.
\item Prof. Dr. Thomas Kneib - for being oral examiner in the disputation.
\item Prof. Dr. Hermann Spellmann - for the decision support to write this thesis and the trustful and cooperative working environment.
\item Christoph Fischer - for valuable help in countless tasks.
\item Dr. Jan Hansen - for the numerous hours of help to connect the TreeGrOSS packages with R.
\item My co-authors Alexander Lange, Elmar Spiegel and Sabine Rumpf - for the successful submissions.
\item Andreas Hill - for many hours with constructive and interesting discussions and for proofreading.
\item Dr. Matthias Schmidt and Dr. Egbert Schönfelder - for supporting statistical and methodological questions.
\item Jan Butschkow, Tim Koddenberg, Michael Hill and Sebastian Ohrmann - for proofreading.
\item The partners from the BioEconomy research project - for supporting practical questions regarding forestry, logistics and wood processing.
\item My family - for miscellaneous supports.
\item Laura - for giving me the courage to tackle the challenges and the strength to carry on.
\end{itemize}