\chapter*{Acknowledgements}

When I started my studies on forest sciences at the University of G{\"o}ttingen in 2006, I would never have imagined that my path would led me to Switzerland finishing a doctoral thesis on forest inventory methods at ETH Zurich. On my entire way, I had the great privilege to meet and be accompanied by many people who supported me, backed me up, shared my passion, sparked my interests on interesting subjects and also became very good friends. I would like to take this opportunity to thank those people from the bottom of my heart and express my deep gratitude.\par 

As having been trained in statistical thinking, I can provide strong further evidence against the hypothesis that the completion of a dissertation thesis is likely without the support of colleagues, friends and family. I particularly want to thank


\renewcommand{\labelitemi}{--}
\begin{itemize}
	\item Prof. Hans Rudolf Heinimann - for giving me the opportunity to continue my work in the field of forest inventory at his chair and giving me the freedom to follow my research interests and explore new ideas.
	
	\item Prof. Joachim Saborowski and Dr. Johannes Breidenbach for acting as co-examiners.
	
	\item Dr. Daniel Mandallaz - for being an excellent and reliable supervisor, for his guidance and always taking the time for questions and discussions, and for benefiting from his great experience in the field of statistics and forest inventory.
	
	\item Alex Massey - for supporting my work with many helpful advice and criticism, and for collaboration and proofreading.
	
	\item Kai Husmann - for his help in many tasks and for countless interesting and constructive discussions concerning our scientific works - and even more for those not concerning it.
	
	\item My colleagues Jochen Breschan, Leo Bont, Daniel Tr{\"u}ssel, Franziska Baumgartner, Maja Messerli, Andreas Gabriel, Sabrina Maurer and the entire group of forest engineering at ETH Zurich for all their support and encouraging words, and for all the nice moments together.
	
	\item The forest service of Rhineland-Palatinate, particularly Dr. Joachim Langshausen, J{\"u}rgen Dietz and Claus-Andreas Lessander, for being a reliable partner, supporting practical questions, and for their openness to new methods and ideas.
	
	\item Peter Kruse for sparking my interest on forestry 13 years ago.
	
	\item My family - my parents Elfie and Joachim, my brother Michael, my grandmother Klara, my parents-in-law Hildegard and Goswin, Andreas and Michael, Martina, Lore and Pauline for all their support and encouragement.
	
	\item Christina - for all her love and support, and having travelled this journey with me side-by-side.	
	
\end{itemize}


\newpage
