	
\chapter{Synthesis}
\label{chap:synth}
\newpage


%---------------------------------------------------------------------------------%
% ---------------------------- Main findings ------------------------------------ %

\section{Main findings}
\label{sec:synth:mainfind}

% ----------------------------------------- %
% * study 3 - estimations:
The investigated design-based small area regression estimators show great potential for the German NFI data to be also (as well) used for estimation of forest attributes on forest district levels. In the case study presented in this thesis, the variance for the estimation of timber volume on two district levels was reduced by 43\% and 25\% on average by the unbiased regression estimators compared to the one-phase estimator exclusively using the terrestrial inventory data. The results thus strongly support the findings of similar studies from other countries that the precision of estimates based on national forest inventory data on regional scales can be considerably improved if the terrestrial inventory data are combined with remote sensing data in the framework of small area estimation techniques. It was also empirically confirmed that the synthetic estimator who neglects the prediction performance of the regression model produces considerably smaller variances and confidence intervals compared to the estimates of the unbiased estimators. In addition to their potential design-bias, this strongly suggests that synthetic estimates are over-optimistic and should be treated with caution. The application of design-unbiased estimators provides the advantage of not having to rely on the validity of the prediction model and should thus be preferred to synthetic estimation whenever it is possible. Fitting the regression model used in the estimators internally seems obvious if the terrestrial inventory phase provides sufficient data for modeling. This is most likely the case if the terrestrial inventory phase is a national forest inventory covering large areas and diverse forest structures. In case the linear regression model is fitted internally, providing the g-weight variance is preferable to the external variance since it partially accounts for the dependency of the regression coefficients on the realized sample and has also better conditional properties as clearly exemplified by the case of post-stratification \citep{mandallaz2016}. One can however always use the external variance formula if other model types such as non-linear models are preferred. However, applying the external variance estimator under models that are fitted internally always leads to underestimating the empirical variance, and this will be more pronounced under small sample sizes. Special attention should be paid to non-parametric models, in which case a bootstrap method has to be applied in order to avoid severe underestimation of the variance \citep{massey2015a, massey2015_thesis}.\par

% , yielding estimation errors of 5\% and 11\% on average respectively.

% ----------------------------------------- %
% * study 2 - modelling:
Whereas calculating the g-weight variance restricts the internal model of the estimators to OLS regression, it was demonstrated that OLS modeling possibilities provide sufficient flexibility to reflect real world dependencies between predictor and response variables and also to mitigate residual inflations caused by quality restrictions in the auxiliary data. If the quality of auxiliary data varies between known strata, including these strata as categorical predictors in the regression model can be an effective means to improve the model precision. Among the auxiliary variables for timber volume prediction, the main tree species of a sample plot revealed to be a powerful predictor when combined with vegetation height information that can be used to boost prediction performance. If categorical data such as tree species information is derived from classification of remote-sensing data and thus prone to classification errors, the calibration technique suggested in this thesis provides a simple and effective method to remove the bias in the regression coefficients caused by misclassifications and thereby increase the model precision. Using categorical variables including their interactions with other categorical or continuous predictors in the regression model thereby extends double-sampling for regression to post-stratification, which is a well-known and effective means to reduce the variance of estimates.\par

% add: vorteil g-weigth bei post-strat.-A gewichte über n1 statt über n2 bei externen varianz

% ----------------------------------------- %
% * study 4 - mapping:
The investigated mapping scenario constitutes a special case of small area estimation where the target area, i.e., a map cell, corresponds to the extent of a sample plot and thus exclusively comprises one model prediction. In the case of prediction maps for continuous response variables, the suggested approach of calculating the user's accuracy for prediction value intervals according to accuracy assessment techniques constitutes an alternative way of providing a confidence interval for each map cell. A difference to classical regression prediction intervals is that the confidence level for the intervals can vary between map cells instead of being fixed. This seems more appropriate for practical usage as it provides the map user with precisions associated to self-defined map value intervals opposed to providing him with map value intervals based on a fixed confidence level. The proposed optimization method can additionally be used to automatically identify prediction intervals that provide highest possible confidence levels while ensuring statistically sound properties of the underlying error matrix.  

% The validation also illustrated that overall goodness-of-fit criteria such as \rmsecv{} or R$^2$ neglects a varying model performance across the prediction space and lead to over- or underestimation of the actual accuracy for specific value ranges of the target variable.

%---------------------------------------------------------------------------------%
% ---------------------------- Limitations and criticism ------------------------ %

\section{Limitations and criticisms}
\label{sec:synth:limits}

% design-based estimators:
The most obvious criticism on design-based estimators is that they are in practice applied to systematic inventory grids although they rely on the assumption of independently and uniformly distributed sample locations. Accounting for the systematic grid design would thus imply the application of model-dependent estimators. Among those, \citet{mandallaz1993} described geostatistical ordinary and double kriging estimators as the most accurate method that provide the advantage of modeling the spatial covariance function. Intensive studies reassuringly revealed that while the design-based estimators tend to overestimate the variance under systematic grids, the difference to the variance under geostatistic kriging is marginal for domains whose spatial extent considerably exceed the range of the spatial correlation. The minimum spatial extent of such domains will more quickly be exceeded in the case of double-sampling, as the spatial correlation range of the model residuals, which in turn contribute most to the double-sampling variance, is usually much smaller than the range of the local density used for one-phase estimations (i.e., ordinary kriging). The overestimation of the variance by design-based estimators under systematic sampling will thus particularly occur for small area estimation, but reassuringly result in conservative confidence intervals.\par

% cluster sampling:
A general criticism on cluster sampling as applied in the German NFI is an increase in the design-based variance compared to simple random sampling, which is however accepted to the benefit of reduced transport costs. However, the inflation in variance for change estimation can be expected to be less than for state estimation. For small area estimation, cluster sampling has the beneficial effect of increasing the terrestrial sample size within the small area domain as it implies an increased probability for at least one plot of a cluster to be included in the small area. However, it has to be emphasized that only part of a cluster being included in a small area domain constitutes an assumption violation of the extended pseudo-synthetic estimator. Whereas this thesis provided empirical evidence that this results in slightly over-optimistic variances, it has also been indicated that the impact is negligible for small area sample size larger than 6. Recommending a minimum small area sample size of 6 for design-unbiased estimation has also been supported by simulation studies for simple random sampling in \citet{mandallaz2013b} which indicated that the nominal coverage rates are ensured for small area sample sizes equal and larger than 6. Re-evaluating the simulation example confirmed the same results for cluster sampling.\par 

% practical implementation:
Another criticism concerns the practical implementation of double-sampling in the special case of small area estimation. A theoretical necessity that was neglected in this thesis is that boundary adjustment on the sample plot and support level should also be performed at the boundaries of a small area unit for both the field and auxiliary data respectively. For the field data, neglecting boundary adjustment can lead to including trees from outside the small area domain in the sampling frame which thus might cause an overestimation of the plot local density. On the part of the auxiliary data, including information from outside the small area domain can weaken the relationship between the derived predictor variable value and the boundary corrected plot local density and thus decrease the model precision \citep{mandallaz2013b}. On the contrary, performing boundary adjustment also at the small area level would imply a considerable increase in effort for data storage and handling, particularly for large scale applications such as presented in this thesis. It could be conjectured that the impacts on estimates are marginal because the occurrence of boundary plots within a domain is usually rare. Since the implications have yet not been investigated, this could be a topic for future work.\par

% modeling:
Regarding the modeling as being a substantial part of regression estimation procedures, time lags between the date of the terrestrial survey and the acquisition of the auxiliary data can constitute a severe limitation. While for state estimation the implied loss in model precision can be partially accounted for by post-stratifying to acquisition dates, it has been shown that such time lags can hamper the application of double sampling for the estimation of change \citep{massey2015_thesis}. Another limiting factor in the modeling framework is angle-count sampling, as it implies the problem that the extraction of the auxiliary data cannot be matched to an exact spatial extent in which the terrestrial data was gathered. Whereas it was confirmed that this can affect model performance, it was revealed that the impacts are not as critical for auxiliary data of medium to low spatial resolution. It could however be conjectured that impacts on model precision are more pronounced when high resolution data are used to make predictions on sample plot level, for example timber volume prediction based on individual tree detection.\par

% mapping:
A further criticism on the presented accuracy assessment for prediction maps is that even the user's accuracies that serve as confidence levels for the prediction value intervals are themselves associated with uncertainties caused by the sampling process of the reference data. Whereas this has also rarely been addressed in similar studies, quantifying these uncertainties is feasible and could be a potential area of future work. In addition, the suggested optimization algorithm is limited by the fact that it constitutes a heuristic approach that rather finds a solution close to the optimum than finding the exact optimal solution. Another implication of the heuristic approach is that the solution for the same input data can slightly vary due to the randomization part of the algorithm and the various possibilities of parametrization.

% - design-based estimators in general:
%  * cluster sampling --> my judgement
%    -> is known to increase the variance of estimates (inflation in variance expressed by cluster correlation coeff.). 
%       Inflation more pronoucend for state than for change estimation (intra cluster correlation for state higher than for change)
%  * comparison of psmall vs. extpsynth:
%    -> psmall and extpsynth performed almost equivalently from a practical point of view, since in all cases they did not yield statist. different point estimates. 
%    -> While for small sample sizes, extpsynth tends to yield marginally smaller variances, its was emprically indicated that the differences between extpsynth and psmall variances are necligible 
%       for sample sizes larger 10 in the small area
%    -> However, extpsynth under cluster sampling -> assumpations can be violated (--> recommendation: for small sample sizes < 6 use psmall)
%  * critical sample sizes n2G in small area --> rule of thumb?
%    -> evidence from case study: a threshold of n2G > 6 seems to avoid problems (viol. of extpsynth assumption, huge increases compared to one-phase variance)
%    -> coverage rates --> simulations (\citep{mandallaz2013b}) also indicate that nominal coverage rates are ensured for sample sizes n2G > 6

%  * design-based estimators rely on randomization of sample locations, but are applied to systematic grids. 
%    -> sorglos wäre man wenn man echt db-based arbeitet (zufällig verteilt) --> aber nicht realistisch aus kostengründen
%    -> db g-weight var. hergeleitet unter lediglicher random annahme. --> unabhängig von der varianz der resdiduen
%    -> tends to overestimate the variance, especially in small areas (safe side -> citation?) --> cite Habil Mandallaz 1993, 2000 
%    -> for large areas, db and geostat very close, especially for two-phase (range of residuals) 
%    -> wald muss als stationär (wald wiederholt sich im Raum) angenommen werden --> scale-dependent! (Bergketten-Beispiel)
%    -> not neglecting would imply the application of model-dependent estimation techniques. Here, geostatistical double-kriging as proposed by Mandallaz (1993) would
%       provide an estimation method that offers a statistically sound incorporation of the spatial correlation & provides method for non-exh. aux.info
%    -> db und geostat stimmen überein für grosse Gebiete, geostat kann besser sein für kleine Gebiete (wenn räumliche Korr.  (range) grösser als Kleingebiet) --> Mandallaz p 151 / 152
%    -> grösster Teil der varianz kommt db aus residuen: --> korrelation der resiuden sehr sehr viel kleiner als der response
%    -> deshalb wenn nur one-phase, ist fehler mit OK deutlich kleiner als db (range response hoch), bei 1-phase kommt es auf den range der residuen an
%        --> range deutlich kleiner, deswegen effekt v.a. in sehr kleinen Gebieten merkbar
%    -> aber geostat. braucht auch Annahmen wie Stationarität 
%    -> mixed-model modellieren abstandunabhängig (innerhalb von Bestand z.B.)

% - practical implementation of double-sampling with special application to SAE:
%  * forest definition (mask vs. field decision) 
%  * boundary adjustment at small area borders
%  * location errors --> refer to lamprecht article --> in the worst case, location errors could lead to falsely exluding or including sample locations and there associated information if they are located close to a domain border
% - practical implementation of double-sampling with special application to SAE: 
%   + forest definition (mask vs. field decision) 
%   + theoretically, boundary adjustment (both on aux. data and field) should be performed also at the boundaries of a SAE unit. For the field inventory, this is at the moment not done, since sae ds estimation has not yet been a implemented part of the National Forest Inventory. (It was consequently also not done for the aux.data in our study). Performing boundary adjustment also at the sa-borders would imply an increased effort in the field inventory as well as an considerable increase in data storage and handling. Neglecting the badj at the sa-borders thus simplify things, but the implications have yet not been investigated. This could be a topic for future work. [When would negclecting be a problem?]. 

% - modelling:
%  * time-lags between terrestrial survey and acquisition of the auxiliary data 
%    -> minimize or avoid time-lags in order to ensure optimal model precision 
%    -> especially important for change estimation
%    -> time-lags could be accounted for in state estimation, but has more limiting effects especially in the case of change estimation (ref to Diss Alex)
%  * angle-count sampling implies the problem that the extraction of the auxiliary information cannot be matched to an exact spatial extent in which the terrestrial data was gathered
%    -> not as critial for the resolution of our auxdata, but likely more severe for higher resolution axdata
%    -> it was shown by Mandallaz (page 240) that 2-3 concentric circles are close to optimal pps (gamma coefficient) --> das könnte man auch für die German NFI machen


% - mapping:
%  * opt. method is a heuristic: 
%    -> output for same input data can slightly vary according to parameterization (cooling scheme) and randomization part  
%    -> no exact solution, rather finds solution close to the optimum
%  * even the user's accuracies that serve as confidence levels for the prediction intervals are themselves associated with uncertainties caused by the sampling process of reference data. Whereas this has also rarely be addressed in 


%  * g-weight variance restricted to OLS regression --> outlook: others models combined with external variance --> hint at Alex Diss
% + multiple-testing: one never knows which of the 95 out of 100 CIs contain the true value


%---------------------------------------------------------------------------------%
% ---------------------------- Implications for future work --------------------- %

\section{Conclusion and implications for future work}
\label{sec:synth:future}

% Main conclusion for each procedure (related to each other):
The results of this thesis strongly confirmed that combining terrestrially gathered forest inventory data with auxiliary information constitutes a very effective means to increase the efficiency of inventories and to expand the range of terrestrial data application. Mapping provides predictions in very high spatial detail and can thus support decisions on very small scales, but on the other side imposes high requirements on the reliability of the predictions as they are purely model-dependent. The findings of this thesis provided evidence that the error associated to specific value intervals are often considerably larger than expected from overall precision accuracies. Conclusively, reliable and detailed information on the map accuracy should always be provided to the practical user. Design-based small area estimators can not meet the spatial detail of model-dependent approaches since they require a minimum number of terrestrial data per small area domain. However, they provide the huge advantage of being robust to model misspecification and are thus very attractive to be used operationally. The recent interest on these family of estimators is also supported by the acceptance of the R-package software that was developed and used in this thesis (4000 downloads since release in June 2016). Both approaches, model-dependent mapping and design-based small area estimation, suggest that one should in any case not discontinue the conduction of terrestrial inventories as they are crucial in order to quantify the reliability of estimates and predictions.\par

% Conclusion and outlook for aux.data and modeling:
This thesis also supports the recommendation to increasingly use already widely available auxiliary data in the service of forest inventories. It should be emphasized that beside high resolution remote sensing data, these auxiliary data can also be provided by maps based on expert judgement (e.g., reflecting soil properties and bioclimatic growing conditions). For change estimation, it could also be beneficial to include information indicating tree species specific growth impacts of intermediate dry events that are likely to occur in increasing frequencies with ongoing climate change. With regards to rapidly increasing availability of more frequently updated and alternative remote sensing data, it can be conjectured that the problem of time lags between terrestrial survey and the auxiliary information will not be an issue in the future. This will also improve the applicability of double-sampling and mapping techniques for the estimation of change. Two challenges of future applications are the reduction of positional inaccuracies of the terrestrial sample locations as well as the synchronization between auxiliary data and terrestrial data derived by angle-count sampling. Concerning the first, various matching approaches to reconstruct the exact plot locations have already been demonstrated. However, the effects of such matching procedures on design-based estimates should first be investigated via simulation studies before being applied operationally. With regards to the problem associated to angle-count sampling, alternatively using 2-3 concentric circles can be very close to optimal probability proportional to size sampling \citep{mandallaz2008} while providing an unambiguous extent in which the auxiliary data can be derived. This should be considered in case angle-count sampling turns out to be a major limiting factor for modeling purposes in future NFIs that are increasingly using multi-phase estimation methods.\par
 
% final conclusion and outlook to geostatistics:
One of the strengths of design-based double-sampling estimation which makes them very attractive to be used operationally is that they can be applied with a high degree of automation in order to frequently produce estimates of numerous forest attributes for a large number of estimation domains with relatively low effort. Developing additional or extending existing open-source software and providing further insights from case studies about the performance of such estimators can further support their application in forest practice. An alternative to design-based estimation that has scarcely been applied in practice is model-dependent geostatistical double-kriging as described in \citet{mandallaz1993}. Whereas it has been shown that double-kriging can provide smaller estimation errors than design-based estimation especially in the case of small area estimation, several issues could possibly limit its practical application. Firstly, the application of kriging estimators necessitate the sophisticated modeling of the spatial covariance function of both the model predictions and residuals. This has to be redone for each application and forest attribute individually and is thus very time-intensive. Secondly, the computation of kriging estimates imposes considerably higher requirements on software and computational power than design-based methods and has thus mostly been restricted to expensive commercial software. However, also open-source geostatistical software has recently been developed in soil sciences \citep{georob2018} that provide the required block-kriging computations and might be suited for double-kriging. Thirdly, modeling the spatial covariance functions requires observations of closely located sample locations in order to appropriately depict the short range correlation. If observations shall be provided by large-scale national forest inventory data that are characterized by long distances of several hundred meters or kilometres between sample locations, the use of cluster sampling is thus inevitable to provide the short distance information. In case all these requirements are fulfilled, it should in addition be considered that double-kriging will only provide a further increase in estimation precision compared to design-based double-sampling if the spatial correlation range is not considerably smaller than the spatial extent of the inventory area \citep{mandallaz1993}. First evaluations of the spatial variograms for the German NFI data that were used in this thesis revealed that the spatial correlation range of the timber volume local density was around 2200 m whereas the range of the predictions and residuals were around 1200 and 200 m respectively. These findings indicate that the application of geostatistical double-kriging estimators might rather be of interest for estimations on very small scales such as forest stands and restricted to few inventory areas. In comparison, the design-based double-sampling procedures provide a time and cost-efficient means to frequently produce numerous estimates for many inventory areas but depend on a minimum number of sample points within the estimation domain.



%Finally, geostatistical double-kriging constitutes an estimation technique which could lead to a further increase in estimation precision but has scarcely been investigated empirically. First evaluations of the spatial variograms for the German NFI data that were used in this thesis revealed that the spatial correlation range of the timber volume local density is around 2200 meters whereas the range of the predictions and residuals are around 1200 and 200 meters respectively. Modelling of the spatial covariance functions is however limited by lack of information for distances smaller than the minimum cluster distance as well as for distances smaller than the within cluster distances. It can thus be concluded that for timber volume estimation, the sampling intensity of national forest inventory data are not sufficient for an appropriate modelling of the spatial covariance function. Cluster sampling 




% - range (RLP data)
% response: 2200 m
% predictions: 1200 m
% residuals: 200 m
%
% -> findings agree to Mandallaz 1993: response shows the highest range (2200 m = between cluster distance), then the predictions show medium range (1200 m)
%    and residuals show considerably smaller spatial correlation range (200 m = approximately the within cluster distance)
% -> problem 1: lack of information for distances smaller than within cluster distances 
% -> problem 2: lack of information for distances smaller than within cluster distances and between custer distances






% - Ausblick auf geostatistic: korrelationstrukturen aus eigener Untersuchung nennen --> es wird nicht viel bringen --> ranges (error in variogramm fitting not solved)
%   read discussion habil daniel
%   * It has however to be mentioned that also the model-dependent geostatistic approach relies on the assumptions of stationarity

% The most obvious recommendation concluded from this thesis is to 

% - recommendation to use ds in forest practice because of the already wide availability of auxiliary data. These auxiliary data can be provided by high resolution remote sensing data, but also
%   by maps based on expert judgment such as soil properties and bioclimatic growing conditions. 
%   ==> my vision: For change estimation, it could also be beneficial to include remote sensing derived information that indicate tree species specific growth impacts of intermediate dry events 
%                  that are likley to occure in increasing frequencies with ongoing climate change.
%   * due to the avail. of auxdata, they should increasingly be used in forest inventory
%   * the problems of time-lags will not arise in the future (recent developments -> satellite repetition rates, ...)
%     ==> this will also improve the applicability of ds techniques for change estimation
%
% - db estimators:
%   * Despite the mentioned limitations, 
%   * The recent interest on the R-package software (4000 downloads up to this day) supports the interest on the application of design-based estimation methods.
%   * While the 
%   * My message: It is the task to further simplify the application of such estimation procedures for forest practice


% - geostatistic


% prefer sae to mapping

% - design-based regression estimator --> good!
% - due to the avail. of auxdata, they should increasingly be used in forest inventory
% - use auxiliary information! The availability will even increase + more frequently updated --> change estimation; + additional attributes
% - we are within a transition period from one-phase to multi-phase inventories. Thus, one should in any case not [verzichten auf] the terrestrial inventory
% - enhance comparability between field data and remote sensing data:
%   * minimize time-lags between terrestrial survey and acquisition of the auxiliary data --> especially crucial for change estimation
%   * improve positional accuracy, spatial resolution
% - use estimators also for:
%   * change estimation
%   * estimation of ratios (--> if desired information is also supported by auxiliary information!)
%
% -> angle count sampling -> Ausblick: it was shown by Mandallaz (page 240) that 2-3 concentric circles are close to optimal pps (gamma coefficient) --> das könnte man auch für die German NFI machen
%
% -> badj: 
%
% - area-wide remote sensing data are more frequently acquired
%   
% Questions:
% - why is sae estimation so attractive? Why should it be applied / is it a MUST to be applied?
% - What recent developments and circumstances facilitate the application of SAE?
%   (frequently availabe and updated remote sensing information and software)
% - what improvements are necessary or can be expected?
% - recommendations?
%
%
% - Ausblick auf geostatistic: korrelationstrukturen aus eigener Untersuchung nennen --> es wird nicht viel bringen --> ranges (error in variogramm fitting not solved)
%   read discussion habil daniel
%   * It has however to be mentioned that also the model-dependent geostatistic approach relies on the assumptions of stationarity






